Marine species do not need visas: the transboundary nature of the world’s exploited marine species

### 2020-04-29 Scripts/Manuscript_Draft.Rmd
Updates for submission

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
# subtitle: "[DRAFT DO NOT SHARE]"
output:
  pdf_document:
    fig_caption: yes
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ./References/FishForVisa_Ref.bib
csl: ./References/nature_cc.csl
---

@@

#### READ ME !!! ####
# Run this chunk before knit so you make sure you have all pkgs installed in R

ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE,repos = "http://cran.us.r-project.org")
  sapply(pkg, require, character.only = TRUE)
}

@@

ipak(packages)

@@

# Abstract
<!-- *An opening sentence that sets the question that you address and is comprehensible to the general reader, background content specific to this study, results, and a concluding sentence. It should be a single paragraph.* -->

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world's identified fished species are transboundary (n = 633) and were responsible for 80% of the global-EEZs catches (48 million tons) and 78.5% of fisheries revenue (USD 65 billion) in the last decade. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that fisheries targeting transboundary species appear worse off than those mainly catching non-transboundary species. Our study has direct implications for the management of transboundary species, pointing to the urgent need for strengthened international cooperation in the management of these fisheries.

# Introduction

@@

# Results and Discussion

@@

![**Figure 1.** Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean +- sd. Revenue in 2010 real USD.](~/GitHub/FishForVisa/Figures/Fig1.png)

@@

![**Figure 2.** Weighted benefits of transboundary species by Km2 and UN sub-region. A) Million 2010 USD. B) Million tons. Point sub-region average bars +- s.d.](~/GitHub/FishForVisa/Figures/Fig2.png)

@@

![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Catchs of main fishing countries fishing each ecosystem preference's top five species in terms of shared EEZs (shown in bold in A) between 2005 and 2014.](~/GitHub/FishForVisa/Figures/Fig4.png)

@@

# Conclusions

@@

# Materials and Methods

## Databases of species' geographic distribution

@@

```{r Data_Table, eval = F, echo = F}

@@

data.frame(Source,
           Abreviation,
           Main_Method,
           Sources) %>% 
  rename("Main Method"=Main_Method) %>% 
  # kable("latex", booktabs = T,
  #       caption = "Summary of data sources to estimate species' distributions. All data has a ressolution of 0.5 degrees latitude x 0.5 degrees longitude. All data is publically available, see references for details.") %>%
  # kable_styling(full_width = F) %>%
  # column_spec(2,width = "8cm") %>% #wraps text in column2
  # column_spec(3,width = "3cm") %>% #wraps text in column2
  write_csv("Table_1.csv")

@@

### *Occurrence data*

@@

### *Distribution models*

@@

### *Catch data*

@@

## Determining transboundary species trait

@@

### Criteria 1; Neighboring EEZs

@@

### Criteria 2; Data agreement

@@

### Criteria 3; Modeling verification

@@

### Criteria 4; Spatial distribution

@@

## Fisheries trends

@@

## Statistical analysis

@@

```{r Status_Table, eval = F, echo = F}

@@

data.frame(Categories,
           Rules) %>% 
  # kable("latex", booktabs = T,
  #       caption = "Rules to determine the category of each transboundary species") %>%
  # kable_styling(full_width = F) %>%
  # column_spec(1,width = "3cm") %>% #wraps text in column2
  # column_spec(2,width = "13cm") %>% #wraps text in column2
  write_csv("Table_2.csv")

@@

## Key uncertainties

@@

Thirdly, we assumed that, if the Sea Around Us reconstructed data recorded a species as caught in any given grid cell, then the species was present within that grid cell. While catch data is not exempt of uncertainty, in most cases, the differences between the Sea Around Us and the FAO self-reported data is smaller towards recent years [@Pauly:2019bf]. Thus, we limited the catch data reference period in our analyses to between 2005 and 2014, the last ten years of data available. Likewise, the spatial allocation of the catch is subjected to impressions, mainly produced by differences in the spatial scale of the original data and the method employed by the Sea Around Us. However, this discrepancy is mainly applicable to data provided by RFMOs in the high seas, which were not included in this study (see http://www.seaaroundus.org/data/#/spatial-catch and @Zeller2016 for mode details in the spatial allocation of the SAU).

@@

# Data Availability
The Sea Around Us datasets used are publicly available at http://www.seaaroundus.org/, the occurrence dataset is all publicly available (see Methods), the ENM-Nereus data can be accessed upon request to G.R. All code is available at https://github.com/jepa/FishForVisa

\clearpage

# References

<div id="refs"></div>

\clearpage

# Acknowledgments
We thank T. Clark and A. Tidd for helpful comments and discussions. G.R. was supported by the EO Wilson foundation. C.W. was supported by the Walton Family Foundation (grant 2018-1371), the David and Lucile Packard Foundation (grant 2019-68336), and the Gordon and Betty Moore Foundation (grant GBMF5668.02). W.W.L.C was supported by the NSERC Discovery Grant and the SSHRC Partnership.

# Authors contributions
J.P.A. Conceptualization, methodology, formal analysis, writing original draft. G.R. Methodology, data curation, writing - review & editing. C.W. Methodology, writing - review & editing. W.W.L.C. Supervision, writing - review & editing. All authors provided valuable input to the study conceptualization and discussed the results and analysis.

# Additional Information  
**Competing interests**: No declared.

@@

img <- readPNG(here("./Figures/FigS4.png"))
grid.raster(img)
```

![Figure S2. Histogram of number of transboundary species using different Area Index threshold values. The darker histogram represents the threshold (25%) used in the present analysis](~/GitHub/FishForVisa/Figures/FigS4.png)

>>>>

---
title: "Marine species do not need visas: the transboundary nature of the world’s exploited marine species"
author: "Authors: Juliano Palacios-Abrantes*^1^, Gabriel Reygondeau^1,2^, Colette Wabnitz^1,3^ and William W.L. Cheung^1^"
date: "Affiliations: ^1^Institute for the Oceans and Fisheries, University of British Columbia, Vancouver, Canada. ^2^Yale University, New Haven, Connecticut, USA. ^3^Stockholm Resilience Center, Stockholm University, Stockholm, Sweden."
subtitle: "[DO NOT SHARE]"
output:
  pdf_document:
    fig_caption: yes
  word_document:
    # reference_docx: ./Manuscript_Instructions/science_template.docx
geometry: margin=1in
editor_options:
  chunk_output_type: inline
bibliography: ../References/FishForVisa_Ref.bib
csl: ../References/nature_cc.csl
---

@@

#### READ ME !!! ####
# Run this chunk before knit so you make sure you have all pkgs installed in R
library(MyFunctions)

@@

my_lib(packages)

@@

## Abstract
<!-- *An opening sentence that sets the question that you address and is comprehensible to the general reader, background content specific to this study, results, and a concluding sentence. It should be a single paragraph.* -->

Regulatory boundaries and species distributions are often not aligned, especially in the case of marine species crossing multiple Economic Exclusive Zones (EEZs). This represents a challenge for fisheries management as policies tend to focus at the national level, yet collaborations are needed to maximize long-term ecological, social and economic benefits. Here, we combined species distributions and the spatial delimitation of EEZs to identify the number of commercially exploited marine species that are shared between neighboring nations. We found that 67% of the world’s identified fished species are transboundary (n = 633) and were responsible for 80% of the global-EEZs catches (48 million tons) and 78.5% of fisheries revenue (USD 65 billion) in the last decade. Some countries received over 90% of their catch and economic benefits from a few shared resources. Our analysis suggests that fisheries targeting transboundary species appear worse off than those mainly catching non-transboundary species. This study has direct implications for the management of transboundary species, pointing to the urgent need for strengthened international cooperation in the management of these fisheries.

## Introduction

@@

## Results and Discussion

@@

![**Figure 1.** Number of transboundary species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by color coding of EEZs and land polygons, respectively. B) Contribution of transboundary species to regional revenue (left) and catch (right). Regions according to the United Nations sub-regions. Points = mean $\mp$ sd. Revenue in 2010 real USD.](~/GitHub/FishForVisa/Figures/Fig1.png)

@@

![**Figure 2.** Weighted benefits of transboundary species by $kmK2$ and UN sub-region. A) Million 2010 USD. B) Million tons. Point sub-region average bars $\mp$ s.d.](~/GitHub/FishForVisa/Figures/Fig2.png)
**NOTE to legend: These are not million. They are probably thousand**

@@

![Figure 4. Number of EEZs shared by transboundary species. A) Number of EEZs by species organized by associated habitat according to FishBase. Showing only species that share > 20 EEZs. B) Catchs of main fishing countries fishing each ecosystem preference’s top five species in terms of shared EEZs (shown in bold in A) between 2005 and 2014](~/GitHub/FishForVisa/Figures/Fig4.png)

@@

## Conclusions

@@

## Materials and Methods

### Databases of species' geographic distribution

@@

```{r Data_Table, eval = T, echo = F}

@@

data.frame(Source,
           Abreviation,
           Main_Method,
           Sources) %>% 
  rename("Main Method"=Main_Method) %>% 
  kable("latex", booktabs = T,
        caption = "Summary of data sources to estimate species' distributions. All data has a ressolution of 0.5 degrees latitude x 0.5 degrees longitude. All data is publically available, see references for details.") %>%
  kable_styling(full_width = F) %>%
  column_spec(2,width = "8cm") %>% #wraps text in column2
  column_spec(3,width = "3cm") #%>% #wraps text in column2
  # write_csv("Table_1.csv")

@@

#### *Occurrence data*

@@

#### *Distribution models*

@@

#### *Catch data*

@@

### Determining transboundary species trait

@@

#### Criteria 1; Neighboring EEZs

@@

#### Criteria 2; Data agreement

@@

#### Criteria 3; Modeling verification

@@

#### Criteria 4; Spatial distribution

@@

### Fisheries trends

@@

### Statistical analysis

@@

```{r Status_Table, eval = T, echo = F}

@@

data.frame(Categories,
           Rules) %>% 
  kable("latex", booktabs = T,
        caption = "Rules to determine the category of each transboundary species") %>%
  kable_styling(full_width = F) %>%
  column_spec(1,width = "3cm") %>% #wraps text in column2
  column_spec(2,width = "13cm") #%>% #wraps text in column2
  # write_csv("Table_2.csv")

@@

### Key uncertainties

@@

Thirdly, we assumed that, if the Sea Around Us reconstructed data recorded a species as caught in any given grid cell, then the species was present within that grid cell. While catch data is not exempt of uncertainty, in most cases, the differences between the Sea Around Us and the FAO self-reported data is smaller towards recent years [@Pauly:2019bf]. Thus, we limited the catch data reference period in our analyses to between 2005 and 2014, the last ten years of data available. Likewise, the spatial allocation of the catch is subjected to impressions, mainly produced by differences in the spatial scale of the original data and the method employed by the Sea Around Us. However, this discrepancy is mainly applicable to data provided by RFMOs in the high seas, which were not included in this study (see http://www.seaaroundus.org/data/#/spatial-catch and [@Zeller2016] for model details in the spatial allocation of the SAU).

@@

<!-- Commented out for Thesis -->
<!-- ## Data Availability -->
<!-- The Sea Around Us datasets used are publicly available at http://www.seaaroundus.org/, the occurrence dataset is all publicly available (see Methods), the ENM-Nereus data can be accessed upon request to G.R. All code is available at https://github.com/jepa/FishForVisa -->

<!-- \clearpage -->

<!-- ## References -->

<!-- <div id="refs"></div> -->

<!-- \clearpage -->

<!-- ## Acknowledgments -->
<!-- We thank T. Clark and A. Tidd for helpful comments and discussions. G.R. was supported by the EO Wilson foundation. C.W. was supported by the Walton Family Foundation (grant 2018-1371), the David and Lucile Packard Foundation (grant 2019-68336), and the Gordon and Betty Moore Foundation (grant GBMF5668.02). W.W.L.C was supported by the NSERC Discovery Grant and the SSHRC Partnership. -->

<!-- ## Authors contributions -->
<!-- J.P.A. Conceptualization, methodology, formal analysis, writing original draft. G.R. Methodology, data curation, writing - review & editing. C.W. Methodology, writing - review & editing. W.W.L.C. Supervision, writing - review & editing. All authors provided valuable input to the study conceptualization and discussed the results and analysis. -->

<!-- ## Additional Information   -->
<!-- **Competing interests**: No declared.  -->

@@

img <- readPNG(here("./Figures/FigS2.png"))
grid.raster(img)
```

![Figure S2. Histogram of number of transboundary species using different Area Index threshold values. The darker histogram represents the threshold (25%) used in the present analysis](~/GitHub/FishForVisa/Figures/FigS2.png)
